### 2023-01-16 docs/chapter_03.qmd
add README and docs

---
title: "Chapter 3"
format: html
---

```{mermaid}
gantt
    dateFormat  YYYY-MM-DD
    title       Capstone implementation timeline

    section Pre-capstone proposal
    Chapter 01            :done,    pre1, 2023-01-10,2023-01-12
    Chapter 02            :active,  pre2, 2023-01-17, 3d
    Chapter 03            :         pre3, 2023-01-19, 2d

    section Data
    Data ingestion pipelines            :datapipe, after pre3, 2023-01-23

    section Modelling
    Establish set of models             :active, m1, after pre3, 1w
    Establish model performance evaluation metrics :active, m2, after a1, 1w
    Implement model training pipelines  :modpipe, after pre3  , 2w
    Train model A                       :after modpipe, 1w
    Train model B                       :after modpipe, 1w
    Train model C                       :after modpipe, 1w
    Evaluate models                     :eval, after modpipe, 4w

    section Capstone paper
    Draft findings                      :draft, after eval, 2w
    Create figures                      :figs, after draft, 1w
    Copyediting and revision            :final, after figs, 2w
```