### 2023-02-02 docs/cpa.qmd
update cpa doc

---
title: "Critical path analysis"
format: html
---

```{mermaid}
gantt
    dateFormat  YYYY-MM-DD
    title      Group Four: Predicting Conflict Escalation

    section Precapstone proposal
    Chapter 01                  :done, pre1, 2023-01-10, 2023-01-12
    Chapter 02                  :done, pre2, 2023-01-17, 3d
    Chapter 03                  :done, pre3, 2023-01-19, 2d

    section Compute and Data
    Cloud infra                 :active, cloud, after pre3, 1w
    Data ingestion pipelines    :done, datapipe, after pre3, 2w
    GDELT full dataset          :active, gdelt, after cloud, 2w
    GDELT sample dataset        :done, gsample, after pre3, 1w
    ACLED dataset               :done, acled, after pre3, 1w
    Sample training dataset     :active, sample, after acled, 1w
    Full training dataset       :dataset, after gdelt, 1w

    section Modelling
    Establish models and approaches to evaluate :done,  m1, after pre3, 1w
    Establish performance evaluation metrics    :active, m2, after a1, 1w
    Model prototyping                           :proto, after dataset, 3w
    Model evaluation                            :eval, after proto, 5w
    Working final model                         :model, after eval, 2w

    section Stretch goals
    API or UI                                   :api, after proto, 2w
    Deployed model                              :prod, after model, 1w

    section Capstone paper
    Draft report                                :draft, after eval, 2w
    Final report                                :report, after draft, 1w
    Presentation                                :pres, after report, 1w
```

```{mermaid}
graph TD
    H[Cloud infra \n\nEST: 1 EFT: 2 \nLST: 3 LFT: 4\n TR: 1 SL: 2] --> A
    A[GDELT \n\nEST: 2 EFT: 4 \nLST: 4 LFT: 6\n TR: 2 SL: 2] -->B(Training dataset \n\nEST: 4 EFT: 5 \nLST: 6 LFT: 7\n TR: 1 SL: 2)
    M[GDELT sample \n\nEST: 1 EFT: 2 \nLST: 2 LFT: 3\n TR: 1 SL: 1] --> I(Sample dataset \n\nEST: 2 EFT: 3 \nLST: 3 LFT: 4\n TR: 1 SL: 1)
    G[ACLED \n\nEST: 1 EFT: 2 \nLST: 2 LFT: 3\n TR: 1 SL: 1] --> B

    G --> I
    B --> C{Working final model    \n\nEST: 11 EFT: 13 \nLST: 12 LFT: 14\n TR: 2 SL: 1}
    B --> J{Model evaluation \n\nEST: 6 EFT: 11 \nLST: 7 LFT: 12\n TR: 5 SL: 1}
    I --> K{Model prototyping \n\nEST: 3 EFT: 6 \nLST: 4 LFT: 7\n TR: 3 SL: 1}

    C --> |Optional|L{Deployed model \n\nEST: 13 EFT: 14 \nLST: 14 LFT: 15\n TR: 1 SL: 1}
    K -->|Optional| D(API or UI \n\nEST: 6 EFT: 8 \nLST: 13 LFT: 15\n TR: 2 SL: 7)

    K --> J
    J --> N(Draft report \n\nEST: 11 EFT: 13 \nLST: 12 LFT: 14\n TR: 2 SL: 1)
    N --> F
    J --> C
    F --> E(Final presentation \n\nEST: 14 EFT: 15 \nLST: 15 LFT: 16\n TR: 1 SL: 1)
    C --> F(Final report \n\nEST: 13 EFT: 14 \nLST: 14 LFT: 15\n TR: 1 SL: 1)
```